---
output:
  pdf_document: default
  html_document: default
---
# A good start is with the data display.
```{r}
data = read.csv("module2_exo5_shuttle.csv",header=T)
data
```

# The data is too small, which will affect the results.
# I think we shouldnt remove the 0 malfuntion

```{r}
logistic_reg = glm(data=data, Malfunction/Count ~ Temperature, weights=Count,family=binomial(link='logit'))
summary(logistic_reg)
```
# I don't like how there is output without a good explanation.

```{r}
plot(data=data, Malfunction/Count ~ Temperature, ylim=c(0,1))
```
# Also, I think the graph from above was enough to understand that there is not a significant impact between temperature and the malfunction.

# Here also  not a very good explanation a little complex.
Suppose that each of the six O-rings is damaged with the same
probability and independently of the others and that this probability
depends only on the temperature. If $p(t)$ is this probability, the
number $D$ of malfunctioning O-rings during a flight at
temperature $t$ follows a binomial law with parameters $n=6$ and
$p=p(t)$. To link $p(t)$ to $t$, we will therefore perform a
logistic regression.

# Why we are doing this if there is no impact between temp and malfunction?
```{r}
# shuttle=shuttle[shuttle$r!=0,] 
tempv = seq(from=30, to=90, by = .5)
rmv <- predict(logistic_reg,list(Temperature=tempv),type="response")
plot(tempv,rmv,type="l",ylim=c(0,1))
points(data=data, Malfunction/Count ~ Temperature)
```


