---
title: "French given names per year per department"
author: "Lucas Mello Schnorr, Jean-Marc Vincent"
date: "October, 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# The problem context
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[https://www.insee.fr/fr/statistiques/2540004](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple


You need to use the _tidyverse_ for this analysis. Unzip the file _dpt2020_txt.zip_ (to get the **dpt2020.csv**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

## Download Raw Data from the website
```{r}
file = "dpt2021_csv.zip"
if(!file.exists(file)){
  download.file("https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2021_csv.zip",
	destfile=file)
}
unzip(file)
```

## Build the Dataframe from file

```{r}
library(tidyverse)

FirstNames <- read_delim("dpt2021.csv",delim=";")
```

All of these following questions may need a preliminary analysis of the data, feel free to present answers and justifications in your own order and structure your report as it should be for a scientific report.
```{r}
# Summary of the data
data_summary <- summary(FirstNames)
print(data_summary)
```

```{r}
head_table <- head(FirstNames,10)
print(head_table)
```
```{r}
install.packages("conflicted")
library(conflicted)

conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

```


1. Choose a firstname and analyse its frequency along time. Compare several firstnames frequency

```{r}
# Load required libraries
library(tidyverse)

# Read data with semicolon delimiter
FirstNames <- read_delim("dpt2021.csv", delim = ";", show_col_types = FALSE)

# Selected names
selected_names <- c("CLÉMENCE", "HÉLOÏSE")

# Filter data for the selected names and valid years 
filtered_data <- FirstNames %>%
  filter(preusuel %in% selected_names) %>%
  filter(str_detect(annais, "^\\d{4}$")) %>%
  mutate(annais_numeric = as.numeric(annais)) %>%
  filter(annais_numeric %% 5 == 0)

# Group and summarize the filtered data
grouped_data <- filtered_data %>%
  group_by(annais, preusuel) %>%
  summarise(total_occurrences = sum(nombre, na.rm = TRUE), .groups = "drop")

# Create a bar chart with bars next to each other for each year
ggplot(grouped_data, aes(x = annais, y = total_occurrences, fill = preusuel)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Occurrences of Selected Names (Every 5 Years)",
       x = "Year",
       y = "Total Occurrences",
       fill = "First Name") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 60, hjust = 1))  # Adjust the size parameter and angle 




```


2. Establish by gender the most given firstname by year. Analyse the evolution of the most frequent firstname.
```{r}

# Remove rows with NA values in the 'annais' column
FirstNames <- FirstNames[complete.cases(FirstNames$annais), ]

# Find the most frequent name for each gender in each year
most_frequent_names <- FirstNames %>%
  group_by(sexe, annais, preusuel) %>%
  summarise(total_occurrences = sum(nombre), .groups = "drop") %>%
  group_by(sexe, annais) %>%
  arrange(desc(total_occurrences)) %>%
  slice(1) %>%
  ungroup()
# Convert 'annais' to numeric
most_frequent_names$annais <- as.numeric(as.character(most_frequent_names$annais))
most_frequent_names <- most_frequent_names %>%
  filter(annais %% 5 == 0)

# Determine the color for each year based on gender
most_frequent_names <- most_frequent_names %>%
  mutate(color = ifelse(sexe == 1, "black", "red"))

# Create a bar chart with facets for each gender
p<-ggplot(most_frequent_names, aes(x = as.factor(annais), y = total_occurrences, fill = color, label = preusuel)) +
  geom_bar(stat = "identity") +
  geom_text(position = position_stack(vjust = 0.1), size = 3, color = "white", angle = 90, hjust = 0) +  # Add labels vertically with white text
  labs(title = "Most Frequent First Name Over Time (Every 5 Years)",
       x = "Year",
       y = "Count") +
  scale_fill_identity() +  
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8, angle = 60, hjust = 1),
        axis.text.y = element_text(size = 6)) +  # Adjust text size for y-axis  
  facet_grid(sexe ~ ., scales = "free_y" ,labeller = labeller(sexe = c("1" = "Men", "2" = "Women"))) +
  theme(legend.position = "none")  # Remove default legend

print(p)
```

3. Optional : Which department has a larger variety of names along time ? Is there some sort of geographical correlation with the data?
