---
output:
  pdf_document: default
  html_document: default
---

###   My computer not working so  I have to work on desktop from university so that why The generating of the random data doesnt exist here but however the data I took it from the software it is the the output of my generating before my laptop crashes.

### I used random data see more precisely the varaibles that impact the value of y ,Note:the generating of the data was in my laptop.
```{r}


data <- read.csv("NewData.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",")


data$Date <- as.POSIXct(data$Date, format = "%Y-%m-%d-%H:%M:%S")

# View the first few rows 
head(data)

```


```{r}
summary(data)
```
### lets try the linear model with the data.

```{r}
# Fit the linear model
model <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11, data = data)

summary(model)

```
###  From above we can see the most important vaiables they are x1,x4,x9,x7 , so lets try to see model with them.
```{r}
# Fit the linear model
model <- lm(y ~ x1 + x4 + x7 + x9 , data = data)

# Check the summary of the model
summary(model)
```
###  Ok we can see how the effect of x7 dispaear on y.
```{r}
par(mfrow = c(2, 2))  # Set up the plotting area to display 4 plots in a 2x2 grid
plot(model)
```

###  Ok there is no pattern but there is a Lot of points not explained by the model.
```{r}
install.packages("car")
library(car)
crPlots(model)
```


###  Here we can see each individual variable, we can see how the X7 is almost a straight line.
```{r}
max_value <- max(data$y, na.rm = TRUE)
max_value
```
```{r}
data_subset <- data[c("x1", "x4","x7","x9", "y")]
data_sorted_desc <- data_subset[order(data_subset$y, decreasing = TRUE), ]
head(data_sorted_desc)
```

###  we can see the highest value for y is 3.225577 , with specific combination of values for the four variables.

###  Now let see another data.

###  Above is random uniform sampling let see if we add to this data my old data that I made it by my hand in the first place.
```{r}

data2 <- read.csv("NewDatawithmyhandData.csv", header = TRUE, stringsAsFactors = FALSE, sep = ",")


data2$Date <- as.POSIXct(data2$Date, format = "%Y-%m-%d-%H:%M:%S")

head(data2)

```

```{r}

# Fit the linear model
model2 <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11, data = data2)

summary(model2)

```

###  This data made add more variable that effect the y let us see if these impact is significant by plotting each variable.

```{r}
# Fit the linear model
model3 <- lm(y ~ x1 + x2 + x3 + x4 + x5 +x7+ x9, data = data2)
summary(model3)
crPlots(model3)
```

###  From above we can see how the just three variable have real impact(x1,x4,x9) and the rest like a straight line, like the model without the hand data.
```{r}
max_value <- max(data2$y, na.rm = TRUE)
max_value
```
###  We get the biggest for y is  3.356876 , now let see the value for each varaible.
```{r}
data_subset <- data2[c("x1", "x4","x7","x9", "y")]
data_sorted_desc2 <- data_subset[order(data_subset$y, decreasing = TRUE), ]
head(data_sorted_desc2)
```

###  We can see the data biased to 0.7 since it was emprical by me additional to the random data.

