---
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading libraries


```{r loading ggplot2}
library(ggplot2)


```


## Loading data For the first experiment


```{r loading csv data into a dataframe}
library(readxl)
Data_Mean_128_16_first <- read_excel("Data_first_experiment_128_16/Data_Mean_128_16_first.xlsx")
Data_128_16_first <- read_excel("Data_first_experiment_128_16/Data_128_16_first.xlsx")


```



## Plotting the data together with the linear regression


```{r}
ggplot(Data_Mean_128_16_first, aes(ID, MT)) +
  geom_point() +
  geom_smooth(method='lm')
```

#### we can see little impact from ID TO MT, Let us go with linear model summary to see more details.


## Linear modelling

```{r linear modeling}
model1 <- lm(MT~ID, data = Data_Mean_128_16_first)
summary(model1)
```



#### Similary to Fitts' law says, ID does seem to have an impact on the movement time. 
#### Also, the R2=0.869 reported by the [experimental software](http://ergo.human.cornell.edu/FittsLaw/FittsLaw.html) and the R2=0.7923 I find here, are different ,  but they are close to each other , but this for mean data not for Raw. 


```{r}
par(mfrow=c(2,2));plot(model1);par(mfrow=c(1,1))

```






#### we can see from the residuLs VS Fitted there s no clear pattern (if i really undersant the pattern from this graph :) ),
#### so we will not reject this model , and we can consider it good , also the pvalue it is inder 0.05 .
#### note this data for the mean not the raw.


#### now ,  same steps but for Raw Data.
```{r }
ggplot(Data_128_16_first, aes(ID, MT)) +
  geom_point() +
  geom_smooth(method='lm')

```




#### Here the we confuse more because the impact doesnt appear that  much  and we have more variablity between the points.

```{r }
model2 <- lm(MT~ID, data = Data_128_16_first)
summary(model2)
```

#### From the summary we can see the R-squared is smaller for Raw data ,that mean there are alot of residuals are not explained by the model.



## Plotting the residuals vs fitted.


```{r}
par(mfrow=c(2,2));plot(model2);par(mfrow=c(1,1))
```

####  From the ploting graph we cant see there is a real impact from ID on MT , but for the  model we can see the fitted vs residuals we cant reject the model and we can see the p value less than 0.05 which is good so we can go with this model ,but the movement values will not just depend on ID , there are factors like the device we use , the age of person , how much he concetrate , does he did somthing else while he doing the experiment , does he use mouse or something else.



## Second Experiment , with same steps from above.

```{r}

library(readxl)
Data_Mean_12_6_second <- read_excel("Data_second_experiment_12_6/Data_Mean_12_6_second.xlsx")
Data_12_6_second <- read_excel("Data_second_experiment_12_6/Data_12_6_second.xlsx")


```

```{r }
ggplot(Data_Mean_12_6_second, aes(ID, MT)) +
  geom_point() +
  geom_smooth(method='lm')

```

#### we cant see the real impact of ID on MT.

```{r}
model3 <- lm(MT~ID, data = Data_Mean_12_6_second)
summary(model3)

```
#### The R-squared bigger than we get from the software 0.640,becuase we reduce the data,still we will not reject the 
#### model,since we have p-value less than 0.05 and there is impact for ID on MT without forgetting the impact for other
#### factors.

```{r}
par(mfrow=c(2,2));plot(model3);par(mfrow=c(1,1))
```




####  Also from this graph residual vs fitted so we cant see the pattern so we cant reject the model , and there is a little impact from the ID on th MT but As I said above there are another factors. 


####  Now for Raw Data,

```{r}

ggplot(Data_12_6_second, aes(ID, MT)) +
  geom_point() +
  geom_smooth(method='lm')

```

```{r }
model4 <- lm(MT~ ID, data = Data_12_6_second)
summary(model4)


```
#### The R-sequared smaller than above (Mean Data)and smaller than we get from the software which was 0.640,
#### Note:maybe because we reduce the number of points so the R-squared wasnt accurate in mean Data, and was bigger.  

```{r }

par(mfrow=c(2,2));plot(model4);par(mfrow=c(1,1))

```

#### We can't reject the model here, there are another factors have impact on the movement related to the human and the device.
#### Note:also there is important thing , which is we don't know how they are calculate the R-squared in their software.